---
title: "Data Exploration"
output: 
  html_document:
    toc: true
    # toc_float: true
---

## Introduction
The purpose of this project is to explore the `colony` and `stressors` datasets that are available on this [website](https://github.com/rfordatascience/tidytuesday/blob/master/data/2022/2022-01-11/readme.md).

## Load required libraries

```{r load-libraries, echo=T, message = FALSE, warning=FALSE}
library(tidyverse)
library(glue)
library(broom)
library(viridis)

```

## Load Data

```{r load-data, message=FALSE}
colony <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-11/colony.csv')

stressor <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-11/stressor.csv')


```


## Inspect Colony Data
```{r}

colony |>
  arrange(months, year) |>
  head() |>
  knitr::kable()

```

```{r}
glimpse(colony)
```



## Clean data
```{r}
colony_df <- colony |>
  filter(!state == "United States") |> #Get all states except 'USA'
  filter(!(year == 2019 & months == 'April-June')) |> # No data collected for this period
  mutate(
    # Put quarters in chronological order
    months = factor(months, levels = c("January-March", "April-June", 
                                       "July-September", "October-December")),
    # colony_max = ifelse(is.na(colony_max), colony_n, colony_max),
    colony_lost_pct = round((colony_lost / colony_max) * 100), 
    # This formula appears consistent with data; colony_lost/colony_max
    colony_added_pct = round((colony_added / colony_max) * 100),
    colony_reno_pct = round((colony_reno / colony_max) * 100)
  ) 

colony_df |>
  head() |>
  knitr::kable()

```


**Inspect Data Again**

```{r}
colony_df |>
  filter(is.na(colony_added)) |>
  head() |>
  knitr::kable()
```


**Get summary statistics for each numeric column**

```{r}
colony_df |>
  select(where(is.numeric)) |>
  summary()

```



## Is there a relationship between colony lost and colony added or colony renovated?

STEPS:  

- Get summary statistics     
- Create label data using summary statistics
- create function to add summary statistics to scatter plot
- see sources below for help



```{r scratch-work, include=FALSE}
# Label Data

# lm_summary <- colony_df |>
#   nest(data = - months) |>
#   mutate(
#     fit = map(data, ~lm(colony_added_pct ~ colony_lost_pct, data = .x)),
#     glance_out = map(fit, broom::glance)
#   ) |>
#   select(months, glance_out) |>
#   unnest(cols = glance_out)
# 
# 
# label_data <- lm_summary |>
#   mutate(
#     r_sqrd = signif(r.squared, 2),
#     p_val = signif(p.value, 2),
#     label = glue("R^2 = {r_sqrd}, p = {p_val}"),
#     colony_lost_pct = 12,
#     colony_added_pct = 105
#   ) |>
#   select(months, label, colony_lost_pct, colony_added_pct)

# label_data

```


```{r summary-statistics}
# Label Data

## function to get linear model estimates
lm_summary_fx <- function(df, x, y, nest_col = months){
  xsym <- rlang::ensym(x) # convert x to a symbol
  ysym <- rlang::ensym(y) # convert y to a symbol
  
  df |>
  nest(data = -{{nest_col}}) |>
  mutate(
    fit = map(data, ~ rlang::inject(lm(!!ysym ~ !!xsym, data = .x))),
    glance_out = map(fit, broom::glance)
  ) |>
  select({{nest_col}}, glance_out) |>
  unnest(cols = glance_out)
}

# lm_summary_fx(colony_df, colony_lost_pct, colony_added_pct, months)

################################################################################
## function to get label data for linear model
label_data_fx <- function(df, x, y, xVal = 12, yVal = 90, facet_by = months){
  
  data <- lm_summary_fx(df, {{x}}, {{y}}, nest_col = {{facet_by}})
  
  data |>
    mutate(
    r_sqrd = signif(r.squared, 2),
    p_val = signif(p.value, 2),
    label = glue::glue("R^2 = {r_sqrd}, p = {p_val}"),
    {{x}} := xVal, # Assign value to x
    {{y}} := yVal # Assign value to y
  ) |>
  select(months, label, {{x}}, {{y}})
}

label_data_fx(colony_df, colony_lost_pct, colony_added_pct) |>
  knitr::kable()

```


```{r plot-colony-function}
library(naniar)

##Function to create plot
## Add R^2 and P-value to plot

plot_colony <- function(df, x, y, xlabel = 'x', ylabel = 'y', ...){
  ggplot(df, aes({{x}}, {{y}})) +
  geom_miss_point() +  #Plot missing variables
  geom_smooth(
    aes(color = 'Regression Line'),
    linewidth = 0.7,
    method = 'lm', formula = y ~ x, na.rm = TRUE
  ) +
  facet_wrap(vars(months)) +
  
  geom_text(
    data = label_data_fx(df, {{x}}, {{y}}, ...),
    # data = label_data,
    aes(label = label),
    size = 9/.pt, # 10 point
    hjust = 0
  ) +
    
    
  scale_x_continuous(
    name = glue::glue('{xlabel}')
  ) +
  scale_y_continuous(
     name = glue::glue('{ylabel}')
  ) +
  scale_color_manual(
    name = '',
    labels = c('Missing Data', 'Available Data', 'Regression Line'), 
    values = c(
      Missing = 'firebrick',
      `Not Missing` = 'springgreen4',
      `Regression Line` = "#000000"
    ),
    guide = guide_legend(override.aes = list(
      linetype = c(0, 0, 1),
      shape = c(19, 19, NA),
      size = c(2, 2, NA)
    ))
  ) +
  theme_bw(12) +
  theme(
    # legend.title = element_blank(),
    legend.position = 'right'
  ) 
}


```


call function `plot_colony`:


```{r colony-lost-vs-colony-added-scatter}
plot_colony(
  df = colony_df, 
  x = colony_lost_pct, 
  y = colony_added_pct, 
  xlabel = 'Colony Lost (%)', 
  ylabel = 'Colony Added (%)',
  yVal = 104
)

```


call function `plot_colony`

```{r colony-lost-vs-colony-reno}
plot_colony(
  df = colony_df, 
  x = colony_lost_pct, 
  y = colony_reno_pct, 
  xlabel = 'Colony Lost (%)', 
  ylabel = 'Colony Renovation (%)',
  xVal = 12,
  yVal = 72
)
```


## Inspect Stressors data


```{r}
# View stressors data

head(stressor) |>
  knitr::kable()

```

**Correct spelling**

```{r stressor-df}
stressor_data <- stressor |>
  mutate(stressor = ifelse(stressor == 'Disesases', 'Diseases', stressor))
```


## What are the top two stressors over the period?


```{r}
stressor_data |>
  filter(!is.na(stress_pct)) |>
  mutate(months = factor(months, 
                         levels = c("January-March", "April-June",
                                    "July-September", "October-December")),
         stressor = fct_reorder(stressor, stress_pct, mean)
  ) |>
  
  ggplot(aes(stress_pct, stressor, fill = stressor)) +
  geom_boxplot() +
  cowplot::theme_minimal_vgrid(11) +
  scale_x_continuous(
    name = 'Stressor (%)'
  ) +
  scale_y_discrete(
    name = ' '
  ) +
  scale_fill_viridis(discrete = TRUE, alpha = 0.6, 
                     option = 'A', direction = -1) +
  theme(
    legend.position = 'none'
  )

```


## How do stressor levels change with time?


```{r}
stressor_data |>
  mutate(months = factor(months, 
                         levels = c("January-March", "April-June",
                                    "July-September", "October-December")),
         stressor = factor(stressor)
  ) |>
  filter(!is.na(stress_pct)) |>
  group_by(year) |>
  ggplot(aes(year, stress_pct, fill = factor(year))) +
  geom_boxplot() +
  viridis::scale_fill_viridis(discrete = TRUE, alpha = 0.6, 
                              begin = 0.05, option = 'A', 
                              direction = -1) +
  theme_bw() +
  scale_x_continuous(
    name = 'Year',
    breaks = 2015:2021
  ) +
  scale_y_continuous(
    name = 'Stressor (%)'
  ) +
  theme(
    legend.position = 'none'
  )

```




```{r}
stressor_data |>
  mutate(months = factor(months, 
                         levels = c("January-March", "April-June",
                                    "July-September", "October-December"))
  ) |>
  filter(!is.na(stress_pct)) |>
  group_by(year, months) |>
  summarise(avg_stress_pct = mean(stress_pct), .groups = 'drop_last') |>
  ggplot(aes(year, avg_stress_pct, color = months)) +
  geom_line(linewidth = 1) +
  theme_classic() +
  ggtitle('Average Stressor vs Time') +
  scale_x_continuous(
    name = 'Year',
    breaks = c(2015:2021)
  ) +
  scale_y_continuous(
    name = 'Average Stressors (%)',
    breaks = seq(7, 16, 2)
  ) +
  theme(plot.title = element_text(hjust = 0.5))

```


## Is there a relationship between colony lost and stressors?

```{r join-colony-stressor}

stressor_df <- stressor_data |>
  mutate(months = factor(months, levels = c("January-March",
                                            "April-June",
                                            "July-September", 
                                            "October-December"))
  ) |>
  filter(!is.na(stress_pct)) |>
  group_by(state, year, months) |>
  summarise(avg_stress_pct = mean(stress_pct), .groups = 'drop_last') |>
  left_join(arrange(.data = colony_df, state, year, months),
                    by = c("state", "year", "months")
  ) |>
  drop_na(colony_lost_pct)
 
stressor_df |>
  head() |>
  knitr::kable()


```

Create a function that will plot the graph.

```{r stressor-scatter-plot-function}
library(naniar)

## Add R^2 and P-value to plot

plot_stressor_fx <- function(df, x, y, xlabel = 'x', ylabel = 'y', yunit = '%', xunit = '%', ...){
  ggplot(df, aes({{x}}, {{y}})) +
  geom_point(aes(color = 'Original Data'), na.rm = T) + 
  geom_smooth(
    aes(color = 'Regression Line'),
    linewidth = 0.7,
    method = 'lm', formula = y ~ x, na.rm = TRUE
  ) +
    
  ggtitle(glue("Relationship between {ylabel} and {xlabel}")) +
    
  facet_wrap(vars(months)) +
  
  geom_text(
    data = label_data_fx(df, {{x}}, {{y}}, ...),
    # data = label_data,
    aes(label = label),
    size = 9/.pt, # 9 point
    hjust = 0
  ) +
    
  scale_x_continuous(
    name = glue::glue('{xlabel} ({xunit})')
  ) +
  scale_y_continuous(
     name = glue::glue('{ylabel} ({yunit})')
  ) +
  scale_color_manual(
    name = '',
    values = c(
      `Original Data` = "#000000", 
      `Regression Line` = 'firebrick'),
    guide = guide_legend(override.aes = list(
      linetype = c(0, 1),
      shape = c(19, NA),
      size = c(2, 0.7)
    ))
  ) +
  theme_bw(12) +
  theme(
    # legend.title = element_blank(),
    legend.position = 'right'
  ) 
}

```


call plot_stressor_fx function...

```{r}
plot_stressor_fx(
  df = stressor_df, 
  x = avg_stress_pct, 
  y = colony_lost_pct, 
  xlabel = 'Average Stress', 
  ylabel = 'Colony Lost',
  xVal = 18,
  yVal = 47 
)
```





## What other information, if any, can we infer from our data? 




SOURCES:  

- https://cran.r-project.org/web/packages/dplyr/vignettes/programming.html  
- https://stackoverflow.com/questions/35023375/r-renaming-passed-columns-in-functions  
- https://stackoverflow.com/questions/65527715/rlang-double-curly-braces-within-lm-formula  